Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-05-06 paper/paper.tex
more stuff

We argue that DMN implication in many
advanced cognitive processes in humans
can be recast as prediction error minimization
based on probabilistic mental simulations,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide blocks of pseudo-experience to iteratively
optimize the behavior of the organism.
%
It has indeed been proposed that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

@@

\begin{wrapfigure}{r}{.4\linewidth}
  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
    the agent takes an action (by following some policy), receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
  \centering
  \includegraphics[width=1.\linewidth]{rl.pdf}
  \label{fig:rl}
\end{wrapfigure}

RL is a problem-solving technique in which,
through repeated interactions with an\textit{environment},
an agent learns to reach goals and optimize reward signals
in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a)$ collected by the agent (Fig. \ref{fig:rl}).
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
an action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

@@

In short, DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic mental
scene simulations coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of MCMC sampling
\citep{tenenbaum2011grow}.

\subsection{Markov decision processes}
We argue that the brain in general, and the DMN in particular,
is a physical system governed by the laws of
physics and can be formally described
by Markov processes at a sufficiently coarse scale.
It has indeed been previously proposed \citep{tegmark2016improved} that
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short.
% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the next state depends not only on the current state
but also on a finite number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include experience episodes of the past.

In artificial intelligence and machine learning, a popular computational model for
multi-step decision processes in such an environment are
\textit{Markov decision processes} (MDPs) \citep{sutton1998reinforcement}.
% For the remainder of this paper, for the sake of simplicity
% we use the abbreviation ``MDP'' to refer to the special case
% of \textit{Partially Observable Markov Decision Processes}.
An MDP operationalizes a sequential decision process
in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., is likely to be different for another agent in the same state).
This is performed alongside each transition in the state space of the environment
 and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities.
This is a minimal set of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}. Model-free RL has had great success in real-world problems including robotics
\citep{ng2004,abbeel2004}, super-human play in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
We propose to use such a model as formal basis to model DMN function.

Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
a core property of human intelligence has been proposed to underly improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
is a rather direct mapping of model-free RL learning algorithms
onto aspects of the brain.
The neurotransmitter dopamine could serve
as a ``teaching signal'' to better estimate value associations
and action policies by controlling
synaptic plasticity in the reward system, including the NAc.
In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space and transitions between states.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
%

In our adopted model-free RL framework,
an agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by a \textit{policy}, which maps states of the world
to probability distributions over actions. Starting at time $t=0$,
followung a policy $\pi$ generates a trajectory of action choices as follows:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0, s_1)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1,s_2)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
    \vdots
  \end{split}
\end{eqnarray*}
We assume time-invariance
in that we expect the dynamics of the process
to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).

@@

The goal of the RL agent is then to update this action policy in order
to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
the definition of cumulative reward $G^\pi$,
the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
viewed to be characteristic for a certain agent.
On the one hand,
setting $\gamma = 0$ yields perfectly hedonistic behavior.
An agent with such a shortsighted horizon is exclusively
concerned with immediate rewards.
This is however not compatible with coordinated planning if long-term goal that is
potentially subserved by neural activity in the DMN.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
On the other hand,
setting $0 < \gamma < 1$ allows a learning process to arise.
A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
that is, the behavioral predispositions related to trading longer delays for higher rewards.
Such an agent puts relatively more emphasis on rewards expected in
a longer-term future. 
Precisely, rewards that are not expected to come within\footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
$\tau := 1/(1 - \gamma)$
time steps from the present point are disregarded. This reduces the variance of expected rewards accumulated across
considered action cascades by limiting the depth of the search tree.
Given that there is more uncertainty in the farsighted future,
it is important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

\subsection{The nuts and bolts of reinforcement learning in the DMN}
Given only the limited information available from an MDP, at a state $s$ the average
utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
% We can now formulate the purpose of the DMN
% as finding a policy $\pi$ for the human agent that maximizes the
% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
\end{equation}
called the $Q$-value for the state-action pair $(s,a)$.
In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
over all considered action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action.
Thus, it answers the question
``What is the expected utility of taking action $a$ in this situation ?''.
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
may well capture processing aspects of the DMN in human agents.

@@

Optimal behavior of the agent corresponds to
a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
% \textbf{XXX please add 1-2 sentences why this is interesting in the current
% context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
  \label{eq:qlearning}
\end{equation}
The learning goal is to approach the
policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
Note that \eqref{eq:qlearning} presents merely a definition and
does not lend itself as a candidate schema
for solving MDPs with even moderately-sized action and state spaces
(i.e., intractability).
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
\begin{equation}
  Q^* = \text{Bel}(Q^*),
  \label{eq:bellman}
\end{equation}
where the so-called Bellman transform $\text{Bel}(Q)$
of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
\begin{equation}
  \begin{split}
   \text{Bel}(Q)(s,a) &:=
   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
%    \; \forall (s,a) \in \mathcal S \times \mathcal A
    % \\&=
    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
    \end{split}
  \end{equation}
The Bellman equation \eqref{eq:bellman} is a temporal consistency equation which provides
a dynamic decomposition of optimal behavior by dividing the $Q$-value function into the immediate
reward and the discounted rewards of the upcoming states.
The optimal Q-value operator $Q^*$ is a fixed point for this equation.
As a consequence of this decomposition, the complicated dynamic programming
problem \eqref{eq:qlearning}
is broken down into simpler sub-problems at different time points.
Exploitation of hierarchical structure in action considerations
has previously been related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a preferred state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Note that in \eqref{eq:bellman} the random sampling
is performed only over quantities which
depend on the environment.
This aspect of the learning process
can unroll off-policy by observing state transitions
triggered by another (possibly stochastic) behavioral policy.

@@

\subsubsection{Optimal control via stochastic gradient descent in the DMN}
Efficient learning of the entire set of model parameters can effectively be achieved
via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
local minima of differentiable functions.
Stochastic here means that the true gradient is estimated from batches of training examples samples from the replay memory, the HC

@@

% \paragraph{Parameter update.}
\begin{equation}
  \delta= -\alpha_k \nabla_{\theta_{k}}\mathcal L(\theta_{k})
  = -\alpha_k\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}],
  \label{eq:oracle}
\end{equation}
where the positive constants $\alpha_1, \alpha_2,\ldots$ are learning rates.
Thus, the next action is taken to drive reward prediction errors
(or \textit{regret}) to percolate from lower to higher processing layers to modulate the
choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
(namely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges) over time (i.e., in the limit $k \rightarrow \infty$)
the thus generated approximating sequence of $Q$-value functions
% \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
$$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$$
are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.

@@

\subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
In RL, Monte-Carlo simulation is a common means to update the agent's belief state
\citep{silver2010monte}.
Monte-Carlo simulation provides a simple method for evaluating the value of a state.
They provide an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
observations and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
These simulations of experience batches
would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
This would be a simple control algorithm by evaluating all legal actions and selecting the action with
highest expected cumulative rewards.
In MDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
\textbf{XXX Elvis: refs needed.}.
Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX Elvis: refs needed in last sentence. Plus, should be really bring up Bayesian?} For sampling-based methods it is still scarce
beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
  activity when humans engage in cognitive processes that are detached from
  the current sensory environment
  \citep{kenet2003spontaneously, fiser2004small, bzdok2013segregation}
  and this network was proposed to be situated
  at the top of the brain network hierarchy
  \citep{carhart2010default, margulies2016situating}.
  Its putative involvement in thinking about the past,
  the future, and hypothetical experiences
  appears to tie in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy encapsulates the repertoire of possible actions
    on the world given a current state.
    The policy matrix encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
% information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  neurobiological implementation of reinforcement learning.}
  Overview of how the constituent nodes of the DMN
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

@@

  From the perspective of our model-free RL agent,
  \textit{inference} in the human brain reduces to
  generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  As such,
  plasticity in the DMN arises naturally:
  if an agent behaving optimally in a given environment moves
  to new, yet unexperienced environments, reward prediction errors will
  massively increase \textbf{Please check/simplify sentence}.
  This will lead to adaptation of the policy until the system converges to a
  new steady-state of optimal action decisions in volatile environments.

>>>>

We argue that DMN implication in many
advanced cognitive processes in humans
can be recast as prediction error minimization
based on probabilistic mental simulations,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain MCMC (MCMC) sampling
\citep{tenenbaum2011grow}.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide blocks of pseudo-experience to iteratively
optimize the behavior of the organism.
%
It has indeed been proposed that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

@@

\begin{wrapfigure}{r}{.4\linewidth}
  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
  \centering
  \includegraphics[width=1.\linewidth]{rl.pdf}
  \label{fig:rl}
\end{wrapfigure}

RL is a problem-solving technique in which,
through repeated interactions with an\textit{environment},
an agent learns to reach goals and optimize reward signals
in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a)$ collected by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

@@

In short, DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic mental
scene of experience simulations coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of MCMC sampling
\citep{tenenbaum2011grow}.

\subsection{Markov decision processes}
Model-free RL has had great success in real-world problems including robotics
\citep{ng2004,abbeel2004},
super-human performance in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
We emphasize that the brain in general, and the DMN in particular,
is a physical system governed by the laws of
physics and can be formally described
by Markov processes at a sufficiently coarse scale.
It has indeed been previously proposed \citep{tegmark2016improved} that
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short.
% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the next state depends not only on the current state
but also on a finite number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include experience episodes of the past.

In artificial intelligence and machine learning, a popular computational model for
multi-step decision processes in such an environment are
\textit{Markov decision processes} (MDPs) \citep{sutton1998reinforcement}.
% For the remainder of this paper, for the sake of simplicity
% we use the abbreviation ``MDP'' to refer to the special case
% of \textit{Partially Observable Markov Decision Processes}.
An MDP operationalizes a sequential decision process
in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., is likely to be different for another agent in the same state).
This is performed alongside each transition in the state space of the environment
 and maintains a probability distribution over the set of possible
states.
% , based on a set of observations and observation probabilities.
This is a minimal set of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}.

Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
a core property of human intelligence has been proposed to underlie improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
is a rather direct mapping of model-free RL learning algorithms
onto aspects of the brain.
The neurotransmitter dopamine could serve
as a ``teaching signal'' to better estimate value associations
and action policies by controlling
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space and transitions between states.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
%

In our adopted model-free RL framework,
an agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by a \textit{policy}, which maps states of the world
to probability distributions over actions.
Starting at time $t=0$,
following a policy $\pi$ generates a trajectory of action choices as follows:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0, s_1)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1,s_2)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
    \vdots
  \end{split}
\end{eqnarray*}
We assume time-invariance
in that we expect the dynamics of the process
to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).

@@

The goal of the RL agent is then to update this action policy in order
to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
the definition of cumulative reward $G^\pi$,
the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
viewed to be characteristic for a certain agent.
On the one hand,
setting $\gamma = 0$ yields perfectly hedonistic behavior.
An agent with such a shortsighted horizon is exclusively
concerned with immediate rewards.
This is however not compatible with coordinated planning of long-term goal that is
potentially subserved by neural activity in the DMN.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
On the other hand,
setting $0 < \gamma < 1$ allows a learning process to arise.
A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
that is, the behavioral predispositions related to trading longer delays for higher rewards.
Such an agent puts relatively more emphasis on rewards expected in
a longer-term future. 
Precisely, rewards that are not expected to come within\footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
$\tau := 1/(1 - \gamma)$
time steps from the present point are disregarded. This reduces the variance of expected rewards accumulated across
considered action cascades by limiting the depth of the search tree.
Given that there is more uncertainty in the farsighted future,
it is important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

\subsection{The components of reinforcement learning in the DMN}
Given only the limited information available from an MDP, at a state $s$ the average
utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
% We can now formulate the purpose of the DMN
% as finding a policy $\pi$ for the human agent that maximizes the
% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
\end{equation}
called the $Q$-value for the state-action pair $(s,a)$.
In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
over all considered action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action.
Thus, it answers the question
``What is the expected utility of taking action $a$ in this situation ?''.
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
may well capture processing aspects of the DMN in human agents.

@@

Optimal behavior of the agent corresponds to
a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
% \textbf{XXX please add 1-2 sentences why this is interesting in the current
% context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
  \label{eq:qlearning}
\end{equation}
The learning goal is to approach the
policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
Note that \eqref{eq:qlearning} presents merely a definition and
does not lend itself as a candidate schema
for solving MDPs with even moderately-sized action and state spaces
(i.e., intractability).
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
\begin{equation}
  Q^* = \text{Bel}(Q^*),
  \label{eq:bellman}
\end{equation}
where the so-called Bellman transform $\text{Bel}(Q)$
of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
\begin{equation}
  \begin{split}
   \text{Bel}(Q)(s,a) &:=
   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
%    \; \forall (s,a) \in \mathcal S \times \mathcal A
    % \\&=
    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
    \end{split}
  \end{equation}
The Bellman equation \eqref{eq:bellman} is a temporal consistency equation which provides
a dynamic decomposition of optimal behavior by dividing the $Q$-value function into the immediate
reward and the discounted rewards of the upcoming states.
The optimal Q-value operator $Q^*$ is a fixed point for this equation.
As a consequence of this decomposition, the complicated dynamic programming
problem \eqref{eq:qlearning}
is broken down into simpler sub-problems at different time points.
Indeed,
exploitation of hierarchical structure in action considerations
has previously been related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a preferred state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Note that in \eqref{eq:bellman} the random sampling
is performed only over quantities which
depend on the environment.
This aspect of the learning process
can unroll off-policy by observing state transitions
triggered by another (possibly stochastic) behavioral policy.

@@

\subsubsection{Optimal control via stochastic gradient descent in the DMN}
Efficient learning of the entire set of model parameters can effectively be achieved
via stochastic \textit{gradient descent}, a common solver algorithm for finding
local minima based on the first derivate of the optimization objective.
Stochastic here means that the true gradient is estimated from batches of training examples samples from the replay memory, the HC

@@

% \paragraph{Parameter update.}
\begin{equation}
  \delta= -\alpha_k \nabla_{\theta_{k}}\mathcal L(\theta_{k})
  = -\alpha_k\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}],
  \label{eq:oracle}
\end{equation}
where the positive constants $\alpha_1, \alpha_2,\ldots$ are learning rates.
Thus, the next action is taken to drive reward prediction errors
to percolate from lower to higher processing layers to modulate the
choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
(namely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges) over time (i.e., in the limit $k \rightarrow \infty$)
the thus generated approximating sequence of $Q$-value functions
% \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
$$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$$
are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.

@@

\subsubsection{Does the hippocampus subserve MCMC sampling?}
In RL, MCMC simulation is a common means to update the agent's belief state
\citep{silver2010monte}.
MCMC simulation provides a simple method for evaluating the value of a state.
They provide an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
observations and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
These simulations of experience batches
would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
This would be a simple control algorithm by evaluating all legal actions and selecting the action with
highest expected cumulative rewards.
In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
\textbf{XXX Elvis: refs needed.}.
Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX Elvis: refs needed in last sentence. Plus, should be really bring up Bayesian?} For sampling-based methods it is still scarce
beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
  activity when humans engage in cognitive processes that are detached from
  the current sensory environment
  and this network was proposed to be situated
  at the top of the brain network hierarchy
  \citep{carhart2010default, margulies2016situating}.
  Its putative involvement in thinking about the past,
  hypothetical experiences, and the future
  appears to tie in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy encapsulates the repertoire of possible actions
    on the world given a current state.
    The policy matrix encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
% information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  neurobiological implementation of reinforcement learning.}
  Overview of how the constituent nodes of the DMN
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

@@

  From the perspective of our model-free RL agent,
  \textit{inference} in the human brain reduces to
  generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  As such,
  plasticity in the DMN arises naturally:
  if an agent behaving optimally in a given environment moves
  to new, yet unexperienced environments, reward prediction errors will
  massively increase \textbf{Please check/simplify sentence}.
  This will lead to adaptation of the policy until the system converges to a
  new steady-state of optimal action decisions in the volatile environment.
